Evaluating the Performance and Energy Efficiency of the \textsc{COSMO-ART} model system

### 2014-04-25 trunk/paper_enahpc14/Sec4:Experimental-results.tex
Manuel's changes

For the experiments with \tinto platform, we also analyze the contribution of the MPI library to the energy consumption and how the use of the blocking versus polling message-passing policies can potentially render energy savings. Specifically, the OpenMPI 1.6.5 library, 
features two operation modes, blocking and polling, which can be selected before \cosmoart
is launched. In the polling configuration (the detault mode) ``MPI'' continually polls
the network interface to check for the completion of an event (e.g. send or receive).
This usually yields low latency but high CPU utilization. Thus, one can expcet that this mode 
attains the best performance, possibly at the cost a higher energy usage. In the blocking
mode, the CPU is yield to other processes/threads if there are no incoming messages.

@@

The TTS and ETS results for a full-day simulation on \tinto are shown in Figure~\ref{fig:4}.  
As expected, the mean energy consumed by COSMO-ART when using the blocking MPI policy is about 
2\,W less than that consumed by the polling mode. This is due to the blocking-waits performed 
by the CPU cores while using the blocking mode, in contrast of the polling-waits that take 
place when polling mode is used. From the point of view of the execution time, there is a 
slight reduction when the blocking MPI policy is leveraged. One can expect that the costs of 
the interruptions used in the blocking MPI policy can lead to an increase of the total 
execution time, however, applications with extensive computing may show better perforamnce with 
the blocking modes. This plays in favour for our results since the energy consumed is also less 
by just changing the behavior of the OpenMPI library from the polling to the blocking policy.

@@

A full tracing experiment is conducted on \tinto in order to capture an overall power
profile at a much finer resolution using the attached PDUs in each node. First we run \cosmoart 
for a full-day simulation and analyze the power consumption. Second we correlate performance 
with power traces of two 1-hor time frames during the day: mid-day (from 11h-12h) and
midnight (23h-24h). All the experiments were performed using 192 \cosmoart processes
mapping 12 processes in each of node.

% For all the experiments we also analyze the contribution of the MPI library to the energy 
% consumption and how the use of the blocking versus polling message-passing policies can 
% potentially render energy savings. Specifically, the OpenMPI 1.6.5 library installed in \tinto, 
% features two operation modes, blocking and modes, which can be selected before \cosmoart is 
% launched. In the polling configuration (the detault mode) ``MPI'' continually polls
% the netwrok interface to check for the completion of an event (e.g. send or receive).
% This usually yields low latency but high CPU utilization. Thus, can one expcet that this mode attanins the best performance, possibley at the cost a higher energy usage. In the blocking
% mode, the CPU is yield to other processes/threads if there are no incoming messages.

In order to obtain an initial overview of the power dissipation of the system model, we have 
obtained the power profiles using our \pmlib framework of a full-day simulation of \cosmoart 
leveraging the polling and blocking MPI policies, these results are depicted in the top/bottom 
plots of Figure~\ref{fig:9}, respectvely. The first observation for 
the polling MPI policy is the quite plain power profile it generates. However, the first and 
last hours of the full-day simulation dissipate slightly less power than the mid-day hours. We 
relate these variations due to the chemical reactions and aerosols, computed by ART, that are 
taking place during the daylight hours but not during night. In this case peak power 
dissipation is about 3821\,W. Looking at the power profile using the blocking MPI policy, we 
can easily observe that the pattern is repeated favoring the reduction of power during 
nighttime hours in a more  noticeable way. The peak power dissipation in this mode is 3768\,W, 
which means a reduction of 50\,W in total or, in other words, 1\,\% of reductuon in the peak 
power. Also, we observe power  drops in each hour simulated, basically related to periods in 
which the processes were performing blocking-waits instead of polling-waits, thus yielding the 
CPU to a low usage during this phases. Our final observation is the reduction in execution time 
when the blocking MPI policy is leveraged in contrast with the polling MPI. One can expect that 
the costs of the interruptions used in the blocking can lead to an increase of the total 
execution time, however, applications with extensive computing may show better perforamnce with 
the blocking mode. This plays in favour for our results since the energy consumed is also less 
by just changing the behavior of the OpenMPI library from the polling to the blocking policy.

To gain more insights about the power-performance behavior of \cosmoart, we also have conducted 
experiments that show power-performance traces of 1 hour simulation. We limited the simulation 
time to this bound since the pattern of the computation performed by \cosmoart is repeated each 
hour, also because the weight of a full-day simulation trace can not be easily handled using 
our machines and memory available (about 45\,GB per trace file). To soften the problem we 
shrink the simulations to 1 hour during mid-day (11h-12h) and midnight (23h-24h), in which 
\cosmoart perform different internal computations, as shown in Figure~\ref{fig:9}. At the same
time we also leverage the two stated MPI policies: polling and waiting modes.

>>>>

For the experiments with \tinto platform, we also analyse the contribution of the MPI library to the energy consumption and how the use of the blocking versus polling message-passing policies can potentially render energy savings. Specifically, the OpenMPI 1.6.5 library, 
features two operation modes, blocking and polling, which can be selected before \cosmoart
is launched. In the polling configuration (the detault mode) ``MPI'' continually polls
the network interface to check for the completion of an event (e.g. send or receive).
This usually yields low latency but high CPU utilization. Thus, one can expcet that this mode 
attains the best performance, possibly at the cost a higher energy usage. In the blocking
mode, the CPU is yield to other processes/threads if there are no incoming messages.

@@

The TTS and ETS results for a full-day simulation on \tinto are shown in Figure~\ref{fig:4}.  
As expected, the mean energy consumed by COSMO-ART when using the blocking MPI policy is about 
2\,W less than that consumed by the polling mode. This is due to the blocking-waits performed 
by the CPU cores while using the blocking mode, in contrast of the polling-waits that take 
place when polling mode is used. From the point of view of the execution time, there is a 
slight reduction when the blocking MPI policy is leveraged. One can expect that the costs of 
the interruptions used in the blocking MPI policy can lead to an increase of the total 
execution time, however, applications with extensive computing, e.g. \cosmoart, may show better perforamnce with the blocking modes This plays in favour for our results since the energy consumed is also less by just changing the behavior of the OpenMPI library from the polling to the blocking policy.

@@

A full tracing experiment is conducted on \tinto in order to capture an overall power
profile at a much finer resolution using the attached PDUs in each node. First we run \cosmoart 
for a full-day simulation and analyse the power consumption. Second we correlate performance 
with power traces of two 1-hor time frames during the day: mid-day (from 11h--12h) and
midnight (23h--24h). All the experiments were performed using 192 \cosmoart processes
mapping 12 processes in each of node.

% For all the experiments we also analyse the contribution of the MPI library to the energy 
% consumption and how the use of the blocking versus polling message-passing policies can 
% potentially render energy savings. Specifically, the OpenMPI 1.6.5 library installed in \tinto, 
% features two operation modes, blocking and modes, which can be selected before \cosmoart is 
% launched. In the polling configuration (the detault mode) ``MPI'' continually polls
% the netwrok interface to check for the completion of an event (e.g. send or receive).
% This usually yields low latency but high CPU utilization. Thus, can one expcet that this mode attanins the best performance, possibley at the cost a higher energy usage. In the blocking
% mode, the CPU is yield to other processes/threads if there are no incoming messages.

In order to obtain an initial overview of the power dissipation of the system model, we have 
obtained the power profiles using our \pmlib framework of a full-day simulation of \cosmoart 
leveraging the polling and blocking MPI policies, these results are depicted in the top/bottom 
plots of Figure~\ref{fig:9}, respectvely. The first observation for 
the polling MPI policy is the quite plain power profile it generates. However, the first and 
last hours of the full-day simulation dissipate slightly less power than the mid-day hours. We 
relate these variations due to the chemical reactions and aerosols, computed by ART, that are 
taking place during the daylight hours but not during night. In this case power consumption
 is about 3821\,W. Looking at the power profile using the blocking MPI policy, we 
can easily observe that the pattern is repeated favoring the reduction of power during 
nighttime hours in a more  noticeable way. The peak power dissipation in this mode is 3768\,W, 
which means a reduction of 50\,W in total or, in other words, 1\,\% of reductuon in the peak 
power. Also, we observe power  drops in each hour simulated, basically related to periods in 
which the processes were performing blocking-waits instead of polling-waits, thus yielding the 
CPU to a low usage during this phases. Our final observation is the reduction in execution time 
when the blocking MPI policy is leveraged in contrast with the polling MPI.

To gain more insights about the power-performance behavior of \cosmoart, we also have conducted 
experiments that show power-performance traces of 1 hour simulation. We limited the simulation 
time to this bound since the pattern of the computation performed by \cosmoart is repeated each 
hour, also because the weight of a full-day simulation trace can not be easily handled using 
our machines and memory available (about 45\,GB per trace file). To soften the problem we 
shrink the simulations to 1 hour during mid-day (11h--12h) and midnight (23h--24h), in which 
\cosmoart perform different internal computations due to the different chemical reactions that happen during the different time frames of the day. At the same time we also leverage the two stated MPI policies: polling and blocking modes. The traces are shown in Figure~\ref{fig:9}.
As can be observed, the repetition
